---
title: "L2"
output: html_document
date: "2023-04-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installing new package

```{r}
install.packages("gitcreds")
library(gitcreds)
```
To check for cred use: gitcreds_set()



Repository 
https://github.com/BuJa96/Pat-L2.git

# Training wheels


creating dummy data
```{r}
library(tidyverse)
wildschwein <- tibble(
    TierID = c(rep("Hans", 5), rep("Klara", 5)),
    DatetimeUTC = rep(as.POSIXct("2015-01-01 00:00:00", tz = "UTC") + 0:4 * 15 * 60, 2)
)

wildschwein
```

calculating time difference

```{r}
wildschwein$diff<- as.numeric(difftime(lead(wildschwein$DatetimeUTC), wildschwein$DatetimeUTC))
```

The problem is now, that the two boars are not being differentiated. the comuptation first nicer. 

```{r}
wildschwein$diff <- mutate(wildschwein, diff = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC)))

wildschwein

wildschwein <- wildschwein |> 
  group_by(TierID) |> 
  mutate(diff = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC)))

```

# Exercise

loading data
```{r}
library("readr") 
library("sf") 

wildschwein_BE <- read_delim("wildschwein_BE_2056.csv", ",")

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE)
```

Now the data is loaded in the format I want. I want to figure out, how many boars there are. 

```{r}
wildschwein_BE$TierID |> 
  unique()
```

This time I want to plot it

```{r}
ggplot(wildschwein_BE, aes(DatetimeUTC, TierName))+
  geom_point()
```

Groupping the boars

```{r}
wildschwein_BE<-wildschwein_BE |> 
  group_by(TierName) |> 
  mutate(diff_s = as.numeric(difftime(lead(DatetimeUTC),DatetimeUTC)))
```

plotting it with various amendments

```{r}
ggplot(wildschwein_BE, aes(diff_s/60))+
  geom_histogram(bindwidth = 1)+
  lims(x=c(0,5000/60))+
  scale_y_log10()
```

going to play some more with the data

```{r}
wildschwein_BE |> 
  filter(DatetimeUTC < "2014-08-24") |>
  ggplot(aes(DatetimeUTC, diff_s, colour=TierName))+
  geom_point()+
  geom_line()
```


Adding some figures to make it easier to calculate the eulidean distance.
Afterwards I want to apply it to the sheet to compute the distance the boars traveled.

```{r}
N1 <- 1204752
N2 <- 1204863
E1 <- 2570409
E2 <- 2570402


wildschwein_BE<- wildschwein_BE |> 
  mutate(steplength_m = sqrt((E-lead(E))^2 + (N-lead(N))^2) )

```

to compute the speed

```{r}
wildschwein_BE<- wildschwein_BE |> 
  mutate(speed_ms = steplength_m/ diff_s)

hist(log10(wildschwein_BE$speed_ms),100)
```

## Task 4

Uploading the new data
```{r}
caro<- read_csv("caro60.csv")
```

Now manually reduce the granularity of our sampling interval by selecting every 3rd, 6th and 9th position and save the output to caro_3, caro_6,caro_9 accordingly.

```{r}

positions_3 <- seq(from = 3, to = length(caro), by = 3)
caro_3 <- caro[positions_3]

positions_6 <- seq(from = 6, to = length(caro), by = 6)
caro_6 <- caro[positions_6]

positions_9 <- seq(from = 9, to = length(caro), by = 9)
caro_9 <- caro[positions_9]

```

### Because the above task is not working, I'll continue with the Caro data. 

Computing Steplengt, timelag and speed
```{r}
#steplength 
caro<-caro|> 
  mutate(steplength_m = sqrt((E-lead(E))^2 + (N-lead(N))^2))

#time lag
caro<-caro |> 
  mutate(diff_s = as.numeric(difftime(lead(DatetimeUTC),DatetimeUTC)))

#speed
caro<- caro |> 
  mutate(speed_ms = steplength_m/ diff_s)
```

Unable to visualize granuality due to error above

```{r}
ggplot(caro)+
  geom_point(aes(E, N), color= caro_3)
```

Task failed... 

# Loading my posmo data

```{r}
posmo <- read_csv("posmo_2023-03-01T00_00_00+01_00-2023-04-28T23_59_59+02_00.csv")
```

Converting it

```{r}
posmo <- st_as_sf(posmo, coords = c("lon_x", "lat_y"), crs = 2056, remove = FALSE)
```

Time difference
```{r}
posmo$diff<- as.numeric(difftime(lead(posmo$datetime), posmo$datetime))
```

Visualize

```{r}
ggplot(posmo)+
  geom_point(aes(transport_mode, diff))

```

